---
title: "MATH1318 Semester 1, 2018"
author: "s3650497- Mohammad,s3689517-Malgorzata Sikora,s3638787 - Ravi Pandey"
subtitle: Time Series - Final Project(Competitve)
output:
  pdf_document: default
  html_notebook: default
  html_document: default
---


```{r include=FALSE}
library(TSA)
library(fUnitRoots)
library(lmtest)
library(readr)
library(forecast)
library(FitAR)
library(forecast)
library(CombMSC)
library(lmtest)
library(fGarch)
library(rugarch)
library(truncnorm)

```


#Importing Data
```{r}

Bitcoin <- read.csv("C:/Users/Wel/Downloads/R/Bitcoin_Historical_Price.csv")
head(Bitcoin)

```
#Converting to Time series Object
```{r}
Bitcoin.ts<- ts(Bitcoin,start=c(2013,4,27),frequency=365.25)
class(Bitcoin.ts)
head(Bitcoin.ts)
plot(Bitcoin.ts,ylab='Price in (USD)',xlab='Year', main = "Time series plot for Bitcoin Price")

```


#Random Walk(Linear Regression Model) 
```{r}

model1 = lm(Bitcoin.ts~time(Bitcoin.ts)) 
summary(model1)
```

#Added the fitted least squares line from model1
```{r}
plot(Bitcoin.ts, ylab='Price in (USD)', main = "Time series plot for simulated linear series")
abline(model1) 

```

#Quadratic Model 
```{r}
t = time(Bitcoin.ts)
t2 = t^2
model1.1 = lm(Bitcoin.ts~t+t2) # label the model as model1
summary(model1.1)

plot(ts(fitted(model1.1)), ylim = c(min(c(fitted(model1.1),
as.vector(Bitcoin.ts))), max(c(fitted(model1.1),as.vector(Bitcoin.ts)))),ylab='Price in (USD)' ,
main = "Fitted quadratic curve to linear data")
lines(as.vector(Bitcoin.ts))




```

#Cyclical or Seasonal Trends
```{r}

Bitcoin.ts1=(ts(Bitcoin.ts, start = 2013,end = 2018, frequency = 12))
month.=season(Bitcoin.ts1) 
model2=lm(Bitcoin.ts1~month.-1)
summary(model2)

model3=lm(Bitcoin.ts1~month.) 
summary(model3)
```

#Cosine Models(Harmonic Models)
```{r}

har.=harmonic(Bitcoin.ts,1) # calculate cos(2*pi*t) and sin(2*pi*t)
model4=lm(Bitcoin.ts~har.)
summary(model4)


##Plotting Harmonic models

plot(ts(fitted(model4),freq=365,start=c(2013,1)),ylab='Price in (USD)',type='l',
ylim=range(c(fitted(model4),Bitcoin.ts)),main="Fitted model to average monthly Bitcoin Price series")
points(Bitcoin.ts)


```

#Stationarity Through Differencing(White noise is generated from the standard normal distribution)

```{r}

plot(Bitcoin.ts,ylab=expression(Y[t]),type='o')

```

#Time series plot of the differences of logarithms of that series
```{r}
par(mfrow=c(2,1))
par(cex=0.7)
plot(Bitcoin.ts, ylab='Price in (USD)',type='l')
first.diff = diff(log(Bitcoin.ts))*100
plot(first.diff,ylab='Change in Log(Price)',type='l')
abline(h=0)

```

#Natural log transformation
```{r}
plot(log(Bitcoin.ts),ylab='Log(Bitcoin)', 
     main = "Time series plot of logarithms of Bitcoin values.")

```

#Differences of the logarithms of the Bitcoin values
```{r}

plot(diff(log(Bitcoin.ts)), ylab='Difference of Log(Bitcoin)', 
     main = "Difference of logarithms for Bitcoin series.")

```
#ADF
```{r}
adf.test(first.diff)
```
#ARMA
Significant Lags can be seen in ACF with first significant lag in PACF. From Eacf, we got ARMA(1,1), ARMA (1,2)
```{r fig.width=12}
#Arma (1,1) (1,2)
par(mfrow=c(1,2))
acf(as.vector(log(Bitcoin.ts)),xaxp=c(0,24,12), main="ACF Bitcoin(USD) Price Time Series.")
pacf(as.vector(log(Bitcoin.ts)),xaxp=c(0,24,12), main=" PACF Bitcoin(USD) Price Time Series.")

eacf(first.diff)
```

#ACF and PACF with absolute values and returns
There are significant lags in both ACF and PACF with deteoriating trend . From Eacf we get ARMA(2,2)& ARMA(2,3)
```{r fig.width=12}
#ARMA(2,2)(2,3)
par(mfrow=c(1,2))
acf(abs(first.diff),main="ACF of the \n
    Absolute Daily Bitcoin Price")
pacf(abs(first.diff),main="PACF of the Absolute Daily Bitcoin Price")


eacf(abs(first.diff))
```


#ACF and PACF for the squared absolute values 
It was squared and we found out that there were no obvious models due to volitality clustering which is more visible for the squared values
there is fuziness due to high variablilty 
```{r fig.width=12}

par(mfrow=c(1,2))
acf(first.diff^2,main="ACF(Squared Daily Bitcoin Price)")
pacf(first.diff^2,main="PACF(Squared Daily Bitcoin Price)")
eacf(first.diff^2)
```

#McLeod Li Test
```{r}
McLeod.Li.test(y=first.diff,main="McLeod-Li Test Statistics for Daily Bitcoin Price")
```
#Shapiro Test
```{r}
qqnorm(first.diff,main="Q-Q Normal Plot of Daily Bitcoin Price")
qqline(first.diff)

shapiro.test(first.diff)
```

#Estimation of Parameters
```{r}

model1 = garch(first.diff,order=c(2,2),trace = FALSE)
summary(model1) # All the coefficients are significant at 5% level of significance.
model1.1 = garchFit(formula = ~garch(2,2), data =first.diff )
summary(model1.1)




```

```{r}

model2 = garch(first.diff,order=c(2,3),trace = FALSE)
summary(model2 )# All the coefficients but aplha_2 are significant at 5% level of significance.
model2.1 = garchFit(formula = ~garch(3,2), data =first.diff, trace = FALSE )
summary(model2.1)

```

```{r, include=FALSE}
residual.analysis <- function(model, std = TRUE,start = 2, class = c("ARIMA","GARCH","ARMA-GARCH")[1]){
  # If you have an output from arima() function use class = "ARIMA"
  # If you have an output from garch() function use class = "GARCH"
  # If you have an output from ugarchfit() function use class = "ARMA-GARCH"
  library(TSA)
  library(FitAR)
  if (class == "ARIMA"){
    if (std == TRUE){
      res.model = rstandard(model)
    }else{
      res.model = residuals(model)
    }
  }else if (class == "GARCH"){
    res.model = model$residuals[start:model$n.used]
  }else if (class == "ARMA-GARCH"){
      res.model = model@fit$residuals
  }else {
    stop("The argument 'class' must be either 'ARIMA' or 'GARCH' ")
  }
  par(mfrow=c(3,3))
  plot(res.model,type='o',ylab='Standardised residuals', main="Time series plot of standardised residuals")
  abline(h=0)
  hist(res.model,main="Histogram of standardised residuals")
  acf(res.model,main="ACF of standardised residuals")
  pacf(res.model,main="PACF of standardised residuals")
  qqnorm(res.model,main="QQ plot of standardised residuals")
  qqline(res.model, col = 2)
  print(shapiro.test(res.model))
  k=0
  LBQPlot(res.model, lag.max = 60, StartLag = k + 1, k = 0, SquaredQ = FALSE)
}

```

#Residual Analysis
The residual analysis for both the models are almost same, by lookin at the time series of the residual analysis we can say that the series is stochastic and without a trend, the histogram is almost normally distributed and the PACF and ACF plots have significant lags and the qq plot for standardised residuals is normally distributed 
```{r fig.height=11, fig.width=12}
residual.analysis(model1,class="GARCH",start=3)

```

```{r fig.height=12, fig.width=12}
residual.analysis(model2,class="GARCH",start=4)

```

#Shapiro Test
```{r}
 shapiro.test(first.diff)
```






#Forecast
```{r}

forecast <- fGarch::predict(model2.1,n.ahead=10,trace=FALSE,plot=TRUE)

```











